### 2015-08-10 Irregular-occ/Irregular-occ.tex
.

\cite{GomesJunior2015} \textcolor{red}{published XXXX precise positions for NNN irregular satellites from 1992-2015} from observations made at the Observatório do Pico dos Dias (OPD), Observatoire Haute-Provence (OHP) and European Southern Observatory (ESO). They showed that the orbits of the irregular satellites of the giant planets have systematic errors. The offsets of the observations relative to the JPL ephemeris could be up to 200 mas for some satellites. These differences could be associated with errors in their orbital elements.

@@

The prediction of the occultations was made by crossing the stellar coordinates and proper motions of the UCAC4 catalogue \citep{Zacharias2013} with the corrected JPL ephemeris as presented in the section \ref{Sec: Offset}. The search for stellar candidates follows the same procedure as presented by \cite{Assafin2010, Assafin2012} and \cite{Camargo2013}.

It was predicted occultation by the 7 major irregular satellites of Jupiter, Phoebe of Saturn and Triton and Nereid of Neptune.

@@

A total of 588 events \textcolor{red}{(Não estou contando ainda os eventos para Tritão)} were identified between January 2015 and December 2017. Table \ref{Tab: occ-list} exemplifies the catalogue of occultations generated and their parameters, which is necessary to produce occultation maps (see Fig. \ref{Fig: ocultacao} as an example). Since these objects are very small, the duration of each event is a few seconds.

\begin{table*}
\caption{\label{Tab: sample-cds} Extract of prediction table. \textcolor{red}{Ainda tenho que fazer a tabela melhor, encaixar. Abaixo é um exemplo apenas para visualização.}}
\begin{centering}
\begin{tabular}{ccccccccccccccccccccccc}
\hline 
%\multicolumn{9}{c}{Himalia} \tabularnewline
\hline
% & \multicolumn{2}{c}{Mean errors} & Nr & Nr &  UCAC4       \tabularnewline
d m Year & h m s & \multicolumn{2}{c}{RA (ICRS) Dec} & C/A & P/A & v & D & R* & $\lambda$ & LST & $\Delta e_{\alpha *}$ & $\Delta e_{\delta}$ & pm & ct & $E_{\alpha *}$ & $E_{\delta}$ & $\mu_{\alpha *}$ & $\mu_{\delta}$ \tabularnewline
\hline
 07 10 2015 & 21 52 34. &  10 53 40.1835 & +08 21  00.489 & 0.465 & 200.31 & 39.31 & 6.25 & 16.6 & 179. & 09:48 & -31.0 & -13.0 & ok & uc &   0  &  0 & -13 &  -7 \tabularnewline
 23 10 2015 & 07 54 50. &  11 05 26.6037 & +07 14  32.030 & 1.017 & 201.33  & 36.11 & 6.12 & 15.5 & 16. & 08:59 & -31.0 & -13.0 & ok & uc &  0  &  0 &  -7 &  -4 \tabularnewline
 31 10 2015 & 00 24 19. &  11 10 56.2964 & +06 42  07.110 &  1.801 &  21.84 & 33.94 &  6.04 & 16.9 &  123. & 08:34 & -31.0 & -13.0 & ok & uc & 0  &  0 & -4 & -13 \tabularnewline
 07 11 2015 &  21 22 53. &  11 16 14.3041 & +06 10  05.194 &  0.306 &  202.36 & 31.33 & 5.94 & 16.2 & 162. & 08:09 & -31.0 & -13.0 & ok & uc & 0 & 0 & 0 & 0 \tabularnewline
 08 11 2015 & 04 27 21. &  11 16 25.7881 & +06 08  55.566 & 1.077 & 202.38 & 31.23 & 5.94 & 16.9 & 55. & 08:08 & -31.0 & -13.0 & ok & uc & 0 & 0 & 9 & -10 \tabularnewline
% 09 11 2015  09 01  4.   11 17 11.6890 +06 04  11.474   11 17 11.7030 +06 04  11.980   0.549   22.45  30.79  5.92 15.9  0.5  0.5  0.5   346. 08:04     -31.0    -13.0 ok uc 0    0    0   19  -13 \tabularnewline
% 11 11 2015  12 58 46.   11 18 33.8286 +05 55  43.695   11 18 33.8357 +05 55  43.946   0.274   22.59  29.98  5.90 14.6  0.4  0.4  0.4   285. 07:57     -31.0    -13.0 ok uc 0    0    0  -14    1 \tabularnewline
% 12 11 2015  20 14 22.   11 19 22.2740 +05 50  40.732   11 19 22.3183 +05 50  42.312   1.713   22.68  29.48  5.88 15.7  0.4  0.4  0.4   175. 07:52     -31.0    -13.0 ok uc 0    0    0   -7   -4 \tabularnewline
% 04 12 2015  03 31 56.   11 30 41.9967 +04 37  36.039   11 30 42.0020 +04 37  36.217   0.196   24.18  20.14  5.56 15.0  0.0  0.0  0.0    47. 06:40     -31.0    -13.0 ok uc 0    0    0    5   -9 \tabularnewline
\hline
\end{tabular}
\par\end{centering}
\textbf{Notes}. Occultation table: day of the year and UTC time of the prediction; right ascension and declination of the occulted star - in the original table, these coordinates are immediately followed by the geocentric astrometric equatorial coordinates (corrected for the offset or ephemeris method), not presented here, of the occulting body; C/A: the geocentric closest approach, in arcseconds; P/A: the planet position angle with respect to the occulted star at C/A, in degrees; velocity in plane of sky, in km s$^{-1}$: positive = prograde, negative = retrograde; D: planet range to Earth, in AU; R*: normalized magnitude to a common shadow velocity of 20 km s$^{-1}$ by the relationship $\textrm{Mag}* = \textrm{Mag}_{\textrm{actual}} + 2.5 \times \log 10 \left(\frac{\textrm{velocity}} {20 \textrm{km}\, \textrm{s}^{-1}} \right)$. A value of 50.0 means that the star is not in the 2MASS; $\lambda$: east longitude of subplanet point in degrees, positive towards east; LST: UT + $\lambda$: local solar time at subplanet point, hh:mm; $\Delta e_{\alpha *}$ and $\Delta e_{\delta}$:offset in mas applied to the ephemeris right ascension and declination, respectively; pm: ok = proper motion applied, no = no proper motion applied; catalogue cross-identification (ct) = uc (UCAC2), 2m (2MASS), fs (field star); f = multiplicity flag (see Table 3); $E_{\alpha *}$ and $E_{\delta}$: uncertainties (mas) in right ascension and declination. A value of 9999 means that there was no estimation of the respective uncertainty; $\mu_{\alpha *}$ and $\mu_{\delta}$: proper motions in right ascension and declination, respectively (mas/year).
\label{Tab: occ-list}
\end{table*}

\begin{figure}
\begin{centering}
\includegraphics[scale=0.2]{figures/Himalia_2015-10-23T07:54:47.png} 
\caption{Occultation map for Himalia \textcolor{red}{Mapa de Himalia para exemplificar, posso colocar outro depois}. Compare the information given in caption with the second entry of Table \ref{Tab: occ-list}}
\label{Fig: ocultacao}
\end{centering}
\end{figure}

@@

Observe a stellar occultation by an irregular satellites demands a great effort. The shadow covers a very restricted area on Earth because the irregular satellites are small. So before we start a large observational campaign, we tested some occultation predictions for larger objects, to assess the quality of the prediction.

This test consists in observing the object and star to be occulted near the event when the two objects are present in the same field of view, preferably when the objects are close to each other. Thus the relative positions between the two objects will have minimal influence of the errors of the catalogue of stars used and possible field distortions \citep[and references therein]{Peng2008}. The offsets of the positions of the star and the satellites will be used to correct the predictions.

To date, two occultation tests were performed, one by Himalia that occurred on March 3, 2015 and the second by Elara that occurred on March 30, 2015. For each event, four maps were generated: the first with the nominal positions of the star and the satellite to the predicted time; the second with the offset calculated as described in section \ref{Sec: Offset}; the third with star and satellite offsets from observations made a few days before the occultation when the two objects were separated (different FOV); and the fourth from observations made when the star and the satellite were close in the same FOV.

Figure \ref{Fig: occ-Himalia}. Shows the four maps for Himalia occultation test in March 03, 2015. The map \ref{Fig: occ-Himalia-nooff} is the nominal prediction with the coordinate of the star given by the catalogue and that of the satellite from the ephemeris. We then corrected the positions of the satellite by an offset calculated by the method showed in section \ref{Sec: Offset} to make the map \ref{Fig: occ-Himalia-ajuste}.

The map \ref{Fig: occ-Himalia-off22fev} was made from obtained positions on February 22 observed with the Zeiss telescope at the Observatório do Pico dos Dias (OPD). On that day, Himalia and the star were observed in separate FOVs as they were still far apart. On the night of the event, March 3, the objects were observed with Perkin-Elmer telescope at OPD just over an hour after the time scheduled for the event. Satellite and star were separated by about 16 arcsec, so very close to each other. From the calculated offsets, the map \ref{Fig: occ-Himalia-off03mar} was generated.

\begin{figure*}
\begin{centering}
\subfigure[Nominal prediction]{\includegraphics[scale=0.19]{figures/Himalia1_2015-03-03T00:39:25.png}  \label{Fig: occ-Himalia-nooff}}
\subfigure[\textcolor{red}{Offset de Himalia a partir do ajuste}]{\includegraphics[scale=0.19]{figures/Himalia2_2015-03-03T00:39:18.png}  \label{Fig: occ-Himalia-ajuste}}
\subfigure[Observation at February 22 (different FOVs)]{\includegraphics[scale=0.19]{figures/Himalia3_2015-03-03T00:39:40.png}   \label{Fig: occ-Himalia-off22fev}}
\subfigure[Observation at March 03 (same FOV)]{\includegraphics[scale=0.19]{figures/Himalia4_2015-03-03T00:39:15.png}   \label{Fig: occ-Himalia-off03mar}}
\caption{Predictions for Himalia: The big red dot show the geocentric closest approach of the shadow. The small red ones are the center of the shadow separated by one minute. The straight lines show the size of the shadow. \ref{Fig: occ-Himalia-nooff} is the map using the nominal positions of the star and satellite. \ref{Fig: occ-Himalia-ajuste} shows the shadow given an estimated offset for the position of Himalia related to the JPL ephemeris obtained in section \ref{Sec: Offset}. In \ref{Fig: occ-Himalia-off22fev} we apply offsets to the positions of star and satellite accordingly the observations made at February 22. In \ref{Fig: occ-Himalia-off03mar} is as in \ref{Fig: occ-Himalia-off22fev} but with observations made at March 03 when the objects were close to each other. \textcolor{red}{Figuras ainda serão mudadas. Colocadas para visualização}}
\label{Fig: occ-Himalia}
\end{centering}
\end{figure*}

In this event, it is possible to see that the shade does not vary much among the four maps suggesting that, at least for Himalia, there is a greater probability of observing an event. In fact, the biggest difference between the shadows of the four maps are 25s and 130km in the direction perpendicular to the shadows (see table \ref{Tab: comparison-Himalia}).

@@

The second test was with the satellite Elara, which is the second biggest irregular satellites of Jupiter. The event was predict the occur at March 30, 2015. The observations were taken on March 25 and April 2, 2015 with the Zeiss telescope. On the night of April 2 they could still be observed in the same FOV. Due to Elara be much weaker, dispersions of the satellite positions on both nights ended up being higher than for Himalia. Still, the differences between the maps obtained were relatively small. The biggest difference between them is 73s and 302km (see table \ref{Tab: comparison-Elara}).

>>>>

\cite{GomesJunior2015} published 6523 precise positions for 18 irregular satellites from observations made at the Observatório do Pico dos Dias (OPD), Observatoire Haute-Provence (OHP) and European Southern Observatory (ESO) between 1992 and 2014. They showed that the orbits of the irregular satellites of the giant planets have systematic errors. The offsets of the observations relative to the JPL ephemeris could be up to 200 mas for some satellites. These differences could be associated with errors in their orbital elements.

@@

The prediction of the occultations was made by crossing the stellar coordinates and proper motions of the UCAC4 catalogue \citep{Zacharias2013} with the corrected JPL ephemeris as presented in section \ref{Sec: Offset}. The search for stellar candidates follows the same procedure as presented by \cite{Assafin2010, Assafin2012} and \cite{Camargo2013}.

We predicted occultation for the 7 major irregular satellites of Jupiter, Phoebe of Saturn and Triton and Nereid of Neptune.

@@

A total of 588 events \textcolor{red}{(Não estou contando ainda os eventos para Tritão)} were identified between January 2015 and December 2017. Table \ref{Tab: occ-list} exemplifies the catalogue of occultations generated and their parameters, which are necessary to produce occultation maps (see Fig. \ref{Fig: ocultacao} as an example). Since these objects are very small, the duration of each event is a few seconds.

\begin{table*}
\caption{\label{Tab: sample-cds} Stellar occultation sampled predictions for Himalia \textcolor{red}{Consertar tabela}}
\begin{centering}
\begin{tabular}{ccccccccccccccccccccccc}
\hline 
%\multicolumn{9}{c}{Himalia} \tabularnewline
\hline
% & \multicolumn{2}{c}{Mean errors} & Nr & Nr &  UCAC4       \tabularnewline
d m Year & h m s & \multicolumn{2}{c}{RA (ICRS) Dec} & C/A & P/A & v & D & R* & $\lambda$ & LST & $\Delta e_{\alpha *}$ & $\Delta e_{\delta}$ & pm & ct & $E_{\alpha *}$ & $E_{\delta}$ & $\mu_{\alpha *}$ & $\mu_{\delta}$ \tabularnewline
\hline
 07 10 2015 & 21 52 34. &  10 53 40.1835 & +08 21  00.489 & 0.465 & 200.31 & 39.31 & 6.25 & 16.6 & 179. & 09:48 & -31.0 & -13.0 & ok & uc &   0  &  0 & -13 &  -7 \tabularnewline
 23 10 2015 & 07 54 50. &  11 05 26.6037 & +07 14  32.030 & 1.017 & 201.33  & 36.11 & 6.12 & 15.5 & 16. & 08:59 & -31.0 & -13.0 & ok & uc &  0  &  0 &  -7 &  -4 \tabularnewline
 31 10 2015 & 00 24 19. &  11 10 56.2964 & +06 42  07.110 &  1.801 &  21.84 & 33.94 &  6.04 & 16.9 &  123. & 08:34 & -31.0 & -13.0 & ok & uc & 0  &  0 & -4 & -13 \tabularnewline
 07 11 2015 &  21 22 53. &  11 16 14.3041 & +06 10  05.194 &  0.306 &  202.36 & 31.33 & 5.94 & 16.2 & 162. & 08:09 & -31.0 & -13.0 & ok & uc & 0 & 0 & 0 & 0 \tabularnewline
 08 11 2015 & 04 27 21. &  11 16 25.7881 & +06 08  55.566 & 1.077 & 202.38 & 31.23 & 5.94 & 16.9 & 55. & 08:08 & -31.0 & -13.0 & ok & uc & 0 & 0 & 9 & -10 \tabularnewline
% 09 11 2015  09 01  4.   11 17 11.6890 +06 04  11.474   11 17 11.7030 +06 04  11.980   0.549   22.45  30.79  5.92 15.9  0.5  0.5  0.5   346. 08:04     -31.0    -13.0 ok uc 0    0    0   19  -13 \tabularnewline
% 11 11 2015  12 58 46.   11 18 33.8286 +05 55  43.695   11 18 33.8357 +05 55  43.946   0.274   22.59  29.98  5.90 14.6  0.4  0.4  0.4   285. 07:57     -31.0    -13.0 ok uc 0    0    0  -14    1 \tabularnewline
% 12 11 2015  20 14 22.   11 19 22.2740 +05 50  40.732   11 19 22.3183 +05 50  42.312   1.713   22.68  29.48  5.88 15.7  0.4  0.4  0.4   175. 07:52     -31.0    -13.0 ok uc 0    0    0   -7   -4 \tabularnewline
% 04 12 2015  03 31 56.   11 30 41.9967 +04 37  36.039   11 30 42.0020 +04 37  36.217   0.196   24.18  20.14  5.56 15.0  0.0  0.0  0.0    47. 06:40     -31.0    -13.0 ok uc 0    0    0    5   -9 \tabularnewline
\hline
\end{tabular}
\par\end{centering}
\textbf{Notes}. Entries included: day of the year and UTC time of the prediction; right ascension and declination of the occulted star - in the original table, these coordinates are immediately followed by the geocentric astrometric equatorial coordinates of the satellite (corrected the computed ephemeris offset using Eqs \ref{eq:linear} or \ref{eq:seno}); C/A: the geocentric closest approach, in arcseconds; P/A: the planet position angle with respect to the occulted star at C/A, in degrees; velocity in plane of sky, in km s$^{-1}$: positive = prograde, negative = retrograde; D: planet range to Earth, in AU; R*: normalized magnitude to a common shadow velocity of 20 km s$^{-1}$ by the relationship $\textrm{Mag}* = \textrm{Mag}_{\textrm{actual}} + 2.5 \times \log 10 \left(\frac{\textrm{velocity}} {20 \textrm{km}\, \textrm{s}^{-1}} \right)$. A value of 50.0 means that the star is not in the 2MASS; $\lambda$: east longitude of subplanet point in degrees, positive towards east; LST: UT + $\lambda$: local solar time at subplanet point, hh:mm; $\Delta e_{\alpha *}$ and $\Delta e_{\delta}$: computed offsets in mas applied to the ephemeris right ascension and declination, respectively; pm: ok = proper motion applied, no = no proper motion applied; catalogue cross-identification (ct) = uc (UCAC2), 2m (2MASS), fs (field star); $E_{\alpha *}$ and $E_{\delta}$: uncertainties (mas) in right ascension and declination. A value of 9999 means that there was no estimation of the respective uncertainty; $\mu_{\alpha *}$ and $\mu_{\delta}$: proper motions in right ascension and declination, respectively (mas/year).
\label{Tab: occ-list}
\end{table*}

\begin{figure}
\begin{centering}
\includegraphics[scale=0.2]{figures/Himalia_2015-10-23T07:54:47.png} 
\caption{Occultation map for Himalia regarding to the second event sampled in Table \ref{Tab: occ-list}}
\label{Fig: ocultacao}
\end{centering}
\end{figure}

\textcolor{red}{Acrescentar um paragrafo falando do GAIA, cujo primeiro release esta previsto para fins de 2016. Falar que seremos capazes de refazer o trabalho de Gomes-Junior 2015 e com posicoes de estrelas GAIA poderemos predizer melhor os eventos de 2017 em diante. Mas por questao de precaucao e de preparo futuro para campanhas, publicamos ate fim de 2017 com o UCAC4.}

@@

Observe a stellar occultation by an irregular satellites demands a great effort. The shadow covers a very restricted area on Earth because the irregular satellites are small. So, before we start a large observational campaign, we tested some occultation predictions for larger objects, to assess the quality of the prediction.

The tests consisted in observing the object and star to be occulted near the event when the two objects were present in the same field of view (FOV), preferably when the objects were close to each other. Thus, the relative positions between the two objects had minimal influence of the errors of the reference catalogue of stars used and possible field distortions \citep[and references therein]{Peng2008}. The offsets of the positions of the star and of the satellites were used to check the original predictions.

Two occultation tests were performed, one by Himalia that occurred on March 3, 2015 and the second by Elara that occurred on March 30, 2015. For each event, four situations were considered: the first with the nominal positions of the star and the satellite to the predicted time; the second with the offset calculated as described in section \ref{Sec: Offset}; the third with star and satellite offsets from observations made a few days before the occultation when the two objects were separated (different FOV); and the fourth from observations made when the star and the satellite were close in the same FOV.

Figure \ref{Fig: occ-Himalia}shows four zoomed in occultation maps corresponding to each of the four situations for Himalia.. The map \ref{Fig: occ-Himalia-nooff} is the nominal prediction with the coordinate of the star given by the catalogue and that of the satellite from the ephemeris. We then corrected the positions of the satellite by an offset calculated by the method showed in section \ref{Sec: Offset} to make the map \ref{Fig: occ-Himalia-ajuste}. The map \ref{Fig: occ-Himalia-off22fev} was made from obtained positions on February 22 observed with the Zeiss telescope \textcolor{red}{dar informacoes do Zeiss, diametro, FOV em arcmin, pixel scale} at the Observatório do Pico dos Dias (OPD). On that day, Himalia and the star were observed in separate FOVs as they were still far apart. On the night of the event, March 3, the objects were observed with Perkin-Elmer telescope \textcolor{red}{dar as mesmas infos, agora para o 1.6m} at OPD just over an hour after the time scheduled for the event. Satellite and star were separated by about 16 arcsec, so very close to each other. From the calculated offsets, the map \ref{Fig: occ-Himalia-off03mar} was generated. Notice that the shadow path was not predicted to cross the OPD. This was not necessary for testing the prediction.

\begin{figure*}
\begin{centering}
\subfigure[Nominal prediction]{\includegraphics[scale=0.19]{figures/Himalia1_2015-03-03T00:39:25.png}  \label{Fig: occ-Himalia-nooff}}
\subfigure[\textcolor{red}{Offset de Himalia a partir do ajuste}]{\includegraphics[scale=0.19]{figures/Himalia2_2015-03-03T00:39:18.png}  \label{Fig: occ-Himalia-ajuste}}
\subfigure[Observation at February 22 (different FOVs)]{\includegraphics[scale=0.19]{figures/Himalia3_2015-03-03T00:39:40.png}   \label{Fig: occ-Himalia-off22fev}}
\subfigure[Observation at March 03 (same FOV)]{\includegraphics[scale=0.19]{figures/Himalia4_2015-03-03T00:39:15.png}   \label{Fig: occ-Himalia-off03mar}}
\caption{Predictions for Himalia: The big red dot show the geocentric closest approach of the shadow. The small red ones are the center of the shadow separated by one minute. The straight lines show the size of the shadow. \ref{Fig: occ-Himalia-nooff} is the map using the nominal positions of the star and satellite. \ref{Fig: occ-Himalia-ajuste} shows the shadow given an estimated offset for the position of Himalia related to the JPL ephemeris obtained in section \ref{Sec: Offset}. In \ref{Fig: occ-Himalia-off22fev} we apply offsets to the positions of star and satellite accordingly the observations made at February 22. \ref{Fig: occ-Himalia-off03mar} is as in \ref{Fig: occ-Himalia-off22fev} but with observations made at March 03 when the objects were close to each other. \textcolor{red}{Figuras ainda serão mudadas. Colocadas para visualização}}
\label{Fig: occ-Himalia}
\end{centering}
\end{figure*}

In this event, it is possible to see that the shade does not vary by much among the four maps suggesting thatthere was a good probability of observing the event. In fact, the biggest differences between the shadows of the four maps were 25s and 130km in the direction perpendicular to the shadows (see table \ref{Tab: comparison-Himalia}).

@@

The second test was with the satellite Elara, which is the second largest irregular satellite of Jupiter. The event was predicted to occur at March 30, 2015. The observations were made on March 25 and April 2, 2015 with the Zeiss telescope. On the night of April 2 they could still be observed in the same FOV. Due to Elara being fainter than Himalia by 2 magnitudes, dispersions of the satellite positions on both nights ended up being higher than for Himalia. Still, the differences between the maps obtained were relatively small. The biggest differences between them were 73s along the shadow path and 302km perpendicular to it (see table \ref{Tab: comparison-Elara}).
